\propertitle

### 2019-04-19 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-18T21:40:56+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\preauthor{\vspace{-0.5\baselineskip}\begin{center}
\normalsize\sffamily%
\lineskip  0.5em}
\postauthor{\par\end{center}}
\predate{\DTMsetdatestyle{mydate}\begin{center}\footnotesize}
\postdate{\end{center}\vspace{-\medskipamount}}

\setfloatadjustment{figure}{\footnotesize}
\captiondelim{\quad}
\captionnamefont{\footnotesize\sffamily%
}
\captiontitlefont{\footnotesize}
\firmlists*
\midsloppy
% handling orphan/widow lines, memman.pdf
% \clubpenalty=10000
% \widowpenalty=10000
% \raggedbottom
% Downes, memman.pdf
\clubpenalty=9996
\widowpenalty=9999
\brokenpenalty=4991
\predisplaypenalty=10000
\postdisplaypenalty=1549
\displaywidowpenalty=1602
\selectlanguage{british}\frenchspacing

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Beginning of document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\firmlists
\begin{document}
\captiondelim{\quad}\captionnamefont{\footnotesize}\captiontitlefont{\footnotesize}
\selectlanguage{british}\frenchspacing
\maketitle

@@

A necessary step of every plausible inference is the assignment of some
probabilistic premisses -- as many as necessary to arrive at our desired
probabilistic conclusions given the evidence. The necessity of this step
was tersely stated by W.~E.~Johnson %(J.~M.~Keynes's supervisor)
a hundred years ago:
\begin{quote}\footnotesize
  Now the axioms of probability enable us to infer any
  probability-conclusion \emph{only} from probability-premisses. In other
  words, the calculus of probability does not enable us to infer any
  probability-value unless we have some probabilities or probability
  relations \emph{given}. Such data cannot be supplied by the
  mathematician. E.g. the rules of arithmetic and the axioms of the
  probability-calculus are utterly impotent to determine, on the supposed
  knowledge that the throw of a coin must yield either head or tail and
  cannot yield both, the probability that it will yield head or that it
  will yield tail. We must assume that the two co-exclusive and
  co-exhaustive possibilities are \emph{equally probable}, before we can
  estimate the probability of either as being a half of certitude.
  \sourceatright{\citep[Appendix on eduction, \sect~5,
    p.~182]{johnson1924}}
\end{quote}

@@

The maximum-entropy principle was proposed first by Jaynes
\parentext{\cite*{jaynes1957,jaynes1963,jaynes1994_r2003};
  \cite{sivia1996_r2006,hobsonetal1973}} as a way of assigning probability
premisses at a level of intermediate depth -- this point is made
especially clear in \textcite[\sect~11.1]{jaynes1994_r2003}.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-04-19T15:23:02+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation constraints%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample constraints}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\preauthor{\vspace{-0.5\baselineskip}\begin{center}
\normalsize\sffamily%
\lineskip  0.5em}
\postauthor{\par\end{center}}
\predate{\DTMsetdatestyle{mydate}\vspace{0.5\baselineskip}%
\begin{center}\footnotesize}
\postdate{\end{center}\vspace{-\medskipamount}}

\setfloatadjustment{figure}{\footnotesize}
\captiondelim{\quad}
\captionnamefont{\footnotesize\sffamily%
}
\captiontitlefont{\footnotesize}
\firmlists*
\midsloppy
% handling orphan/widow lines, memman.pdf
% \clubpenalty=10000
% \widowpenalty=10000
% \raggedbottom
% Downes, memman.pdf
\clubpenalty=9996
\widowpenalty=9999
\brokenpenalty=4991
\predisplaypenalty=10000
\postdisplaypenalty=1549
\displaywidowpenalty=1602
\raggedbottom
\selectlanguage{british}\frenchspacing

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Beginning of document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\firmlists
\raggedbottom
\begin{document}
\captiondelim{\quad}\captionnamefont{\footnotesize}\captiontitlefont{\footnotesize}
\selectlanguage{british}\frenchspacing
\maketitle

@@

A necessary step of every plausible inference is the assignment of some
probabilistic premisses -- as many as necessary to arrive at our desired
probabilistic conclusions given the evidence. The necessity of this step
was tersely stated by W.~E.~Johnson %(J.~M.~Keynes's supervisor)
a hundred years ago:\vspace{\topsep}
\begin{quote}\footnotesize
  Now the axioms of probability enable us to infer any
  probability-conclusion \emph{only} from probability-premisses. In other
  words, the calculus of probability does not enable us to infer any
  probability-value unless we have some probabilities or probability
  relations \emph{given}. Such data cannot be supplied by the
  mathematician. E.g. the rules of arithmetic and the axioms of the
  probability-calculus are utterly impotent to determine, on the supposed
  knowledge that the throw of a coin must yield either head or tail and
  cannot yield both, the probability that it will yield head or that it
  will yield tail. We must assume that the two co-exclusive and
  co-exhaustive possibilities are \emph{equally probable}, before we can
  estimate the probability of either as being a half of certitude.
  \sourceatright{\citep[Appendix on eduction, \sect~5,
    p.~182]{johnson1924}}
\end{quote}

@@

Here's an example. Suppose that $n$ neurons have been recorded in a
sequence of time bins. We're given the time averages of the numbers of
active neurons and neuron pairs in the sequence. We don't know the exact
sequence of activities or the number $T$ of time bins, although we know
that the latter is large. We want to forecast how many neurons will be
active in a \emph{new} time bin.

In a deep approach to this inference problem we assign the probability for
the possible value of $T$, and the probabilities for several hypotheses
about the sequence of activities in $T+1$ or more time bins. One hypothesis
could be, for example, that the $n$ neurons are a sample of a larger
population of $N$ neurons and for this population there are some sufficient
statistics, out of biological reasons. We could of course deepen this
approach further, with no end, down to assigning the probabilities for the
constitution and motion of the molecules of the $N$ neurons and their
environment, the probabilities for the position and response of the
recording electrodes, and so on. From all these probabilities and the data
we could finally calculate the probability for the activity in the new time
bin.

At the other end, the shallowest approach to this inference problem is to
directly assign the probabilities for the number of active neurons in the
new time bin, somehow using the data we were given. (It must be remarked
that such assignment partly constrains, in a backward fashion, all assignments we
would do at a deeper level.)

The maximum-entropy principle was proposed by Jaynes
\parentext{\cite*{jaynes1957,jaynes1963,jaynes1994_r2003};
  \cite{sivia1996_r2006,hobsonetal1973}} as a way of assigning probability
premisses at a more direct level, as in the approach just described. This
point is made especially clear in \textcite{jaynes1994_r2003},
\sect~11.1:\vspace{0.5\topsep}
\begin{quote}\footnotesize
  If we knew the numbers $N$, $L$, $W$, then this could be solved by direct
  application of Bayes' theorem; without that information, we could still
  introduce the unknowns $N$, $L$, $W$ as nuisance parameters and use
  Bayes' theorem, eliminating them at the end. \textelp{} However, the
  Bayesian solution would not really address our problem; it only transfers
  it to the problem of assigning priors to $N$, $L$, $W$, leaving us back
  in essentially the same situation; how do we assign informative
  probabilities?
\end{quote}
Its idea is to translate \emph{all} data we're given as constraints for our
probability assignment, resolving the leftover freedom by choosing the
probability distribution having highest Shannon entropy, as a measure of
its uncertainty or broadness. It can be shown that the use of this
principle is approximately equivalent, under specific conditions, to a
probability assignment at a deeper level based on an assumption of
exchangeability together with a multinomial probability distribution for
long-run frequencies, or with an assumption about sufficient statistics
\citep{jaynes1986d_r1996,rodriguez1989,rodriguez2002,skilling1989b,skilling1990}[\sect~4.5]{bernardoetal1994_r2000}{diaconisetal1981,portamana2009,portamana2017}.

Owing to its directness, the maximum-entropy principle is straightforward
to use -- and to misuse. In our previous example, for instance, it gives
silly results unless the number of time bins is much larger than the
possible activity values (it has thus been variously misused in the
neuroscientific literature \mynote{refs, especially to Tkacik}).

In the present work we show how to use the maximum-entropy principle to...
